---
title: "4810_Project"
author: "Nandar May"
date: "2024-03-19"
output: html_document
---


# Read and Introduce our Dataset
```{r}
star = read.csv('..//data//starbucks_drinkMenu_expanded.csv')
head(star)
```
# Data Structure
```{r}
str(star)
```



# Convert column Total.Fat..g. / Vitamin.A....DV./ Vitamin.C....DV. / Calcium....DV./ Iron....DV./Caffeine..mg.  
# from chr to num
```{r}
# Replace all NA values with 0 in the entire dataframe
star[is.na(star)] <- 0
star$Total.Fat..g.[238] = 3.2
star$Total.Fat..g. = as.numeric(star$Total.Fat..g.)
star$Vitamin.A....DV. = as.numeric(sub("%", "", star$Vitamin.A....DV.))
star$Vitamin.C....DV. = as.numeric(sub("%", "", star$Vitamin.C....DV.))
star$Calcium....DV. = as.numeric(sub("%", "", star$Calcium....DV.))
star$Iron....DV. = as.numeric(sub("%", "", star$Iron....DV.))
star$Caffeine..mg. = as.numeric(sub("%", "", star$Caffeine..mg.))
star[is.na(star)] <- 0
```


# get all the scatter plot of numerical variables

```{r}
numeric_vars <- sapply(star, is.numeric)
pairs(star[, numeric_vars])

```


# according to drink type, combine some levels of the Beverage category

```{r}
Frappuccino = c('Frappuccino® Blended Coffee','Frappuccino® Blended Crème','Frappuccino® Light Blended Coffee')
Espresso = c('Classic Espresso Drinks','Coffee','Signature Espresso Drinks')
# others: Shaken Iced Beverages, Tazo® Tea Drinks, Smoothies 
star$Drink.Type = ifelse(star$Beverage_category %in% Frappuccino,'Frappuccino',ifelse(star$Beverage_category %in% Espresso,'Espresso','non-coffee'))
```


# from literature review, we omit sugar , Dietary.Fibre..g., trans fat, saturated fat, because Total.Fat..g. and Total.Carbohydrates..g. can represent them

# Model selection
```{r}
library(MASS)
MASS::stepAIC(lm(Calories~Total.Fat..g. + 
     Sodium..mg. + Total.Carbohydrates..g. + 
    Cholesterol..mg.  + Protein..g. + 
    Vitamin.A....DV. + Vitamin.C....DV. + Calcium....DV. + Iron....DV. + 
    Caffeine..mg. + Drink.Type,data=star),direction="both")
```
# from the stepAIC, we get model fit1, Caffeine and Vitamin A were removed
```{r}
fit1 = lm(formula = Calories ~ Total.Fat..g. + Sodium..mg. + Total.Carbohydrates..g. + 
    Cholesterol..mg. + Protein..g. + Vitamin.C....DV. + Calcium....DV. + 
    Iron....DV. + Drink.Type, data = star)

summary(fit1)

library(car)
car::vif(fit1)
```

# Detected multicollinearity between several variables​

# Total Fat and Total Carbohydrates are essential to our model​

# => Removed Protein and Calcium from the model

# fit2 , model without Calcium....DV. , Protein..g.
```{r}
fit2 = lm(formula = Calories ~ Total.Fat..g. + Sodium..mg. + Total.Carbohydrates..g. + 
    Cholesterol..mg. + Vitamin.C....DV.  + 
    Iron....DV. + Drink.Type, data = star)

summary(fit2)

library(car)
car::vif(fit2)
```

# compare fit1, fit2
```{r}
library(lmtest)
lmtest::lrtest(fit2,fit1)
```

```{r}
Anova(fit2)
```

fit3 , model without Sodium..mg.
```{r}
fit3 = lm(formula = Calories ~ Total.Fat..g.  + Total.Carbohydrates..g. + 
    Cholesterol..mg. + Vitamin.C....DV.  + 
    Iron....DV. + Drink.Type, data = star)

summary(fit3)
```

```{r}
Anova(fit3)
```

# we know that different drink type may have different ingredients. For example, smoothies may have more sugar than classic expresso coffee.
# so we want to know is there interaction between Drink.Type and other explanatory variables

```{r}
var_names = c('Protein..g.', 'Total.Fat..g.'  , 'Total.Carbohydrates..g.' , 'Cholesterol..mg.', 'Vitamin.C....DV.'  , 'Iron....DV.')
for (variable_name in var_names) {
  p = ggplot(star, aes(x = star[[variable_name]], y = Calories, color = Drink.Type)) +
    geom_point(aes(color = Drink.Type)) +  # Color points by Drink.Type
    # geom_smooth(method = "lm", se = FALSE) + 
    labs(x = variable_name, y = "Calories", color = "Drink Type")
  print(p)

}

```


# from Scatter plot above, we can find that there seems interaction between Drink type and Total.carbonhydrates; 

#fit4 ,model with interaction between Drink type and Total.carbonhydrates;
```{r}
fit4 = lm(formula = Calories ~ Total.Fat..g.  + Total.Carbohydrates..g. + 
    Cholesterol..mg. + Vitamin.C....DV.  + 
    Iron....DV. + Drink.Type +Drink.Type:Total.Carbohydrates..g., data = star)

summary(fit4)

Anova(fit4)
```
H0: fit3 better, beta_interaction = 0
Ha: fit4 better, beta_interaction != 0
```{r}
lrtest(fit3,fit4)
```
## since lrtest p-value < 0.05,we reject H0, and conclude that model with interaction between Drink type and Total.carbonhydrates is better.
## However, R2 adjusted only increased by 0.32%​ Since simple model is preferred. We stay with Model 3 as final model.


### Residuals 
```{r}
s = summary(fit3)$sigma
plot(fit3$fitted.values, fit3$residuals)
abline(h = 0, col = "red")
abline(h = c(-2, 2)*s, col = "red")
```


```{r}
hist(fit3$residuals)
```

```{r}
qqnorm(fit3$residuals)
qqline(fit3$residuals,col='red')
```

