---
title: "4810_Project"
author: "Nandar May"
date: "2024-03-19"
output: html_document
---

#Questions:

```{r}
star = read.csv('C:\\Users\\wudan\\Downloads\\archive\\starbucks_drinkMenu_expanded.csv')
```


```{r}
numeric_vars <- sapply(star, is.numeric)
pairs(star[, numeric_vars])

```

```{r}
library(MASS)
MASS::stepAIC(lm(Calories~.,data=star),direction="both")
```


## 2. the effect of recycling on Carbon Emission? -- model 2 = model 1 + recycling (create a category )
    a. we add this variable to model in question 1, use anova test to check if it is significant enough?
    b. then we use new model to predict, and compare the prediction results with model1 and model2
    
## 3. Model selection based on model 2, -- model final
    a. MASS::stepAIC(model2,scope= [upper,lower],direction=both)
    b. partial residual analysis on some variable, then do transformation
    c. check outliners and influential points
    4. check multicollinearity, cor, VIF() > 10
    5. use test data to do prediction and check the prediction accuracy  
    

## read csv, create new column <Total.Recycling> how many different kind of trash the people recycle {0,1,2,3,4}
```{r}
set.seed(123)
carbon <- read.csv("..\\data\\Carbon Emission.csv", header = T)

# get the length of list-like string in Recycling, 
List.Recycling <- lapply(carbon$Recycling, function(x) as.list(unlist(strsplit(gsub("\\[|\\]", "", x), ",")))) # i delete the as.numeric function, it still works, I don't know why we had that function in the beginning
lengths <- sapply(List.Recycling, length)
carbon$Total.Recycling <- lengths

```


## 1. create a variable total.gabage = bag.size * bag.count
small : 25, median:45,large:74,extra large:90
```{r}
bag.size <- ifelse(carbon$Waste.Bag.Size == "small", 25,
                   ifelse(carbon$Waste.Bag.Size == "medium", 45,
                          ifelse(carbon$Waste.Bag.Size == "large", 74, 90)))
carbon$Total.Waste <- carbon$Waste.Bag.Weekly.Count * bag.size
```


## split data into train and test data
```{r}
split <- sample(c(TRUE, FALSE), nrow(carbon), replace=TRUE, prob=c(0.7, 0.3))
train <- carbon[split, ]
test <- carbon[!split, ]
```




    a. scatter plot
    b. lm model, F test
    c. interpret each coefficient
    d. use waste bag size and Waste bag count weekly to predict people's carbon emission (use test data)


```{r}
plot(train$Total.Waste, train$CarbonEmission)
waste.fit <- lm(CarbonEmission ~ Vehicle.Monthly.Distance.Km + Transport, data = train)
summary(waste.fit)

plot(waste.fit$fitted.values, waste.fit$residuals)
s = summary(waste.fit)$sigma
abline(h = c(2,-2)*s, line = 'red')

waste.fit <- lm(CarbonEmission ~ Total.Waste, data = train)
summary(waste.fit)
```

```{r}
waste.fit2 <- lm(CarbonEmission ~ Vehicle.Monthly.Distance.Km*Transport, data = train)
summary(waste.fit2)
```
### model comparision, with and without interaction
```{r}
anova(waste.fit,waste.fit2)
```
## 3. Model selection based on model 2, -- model final
    a. MASS::stepAIC(model2,scope= [upper,lower],direction=both)
    b. partial residual analysis on some variable, then do transformation
    c. check outliners and influential points
    4. check multicollinearity, cor, VIF() > 10
    5. use test data to do prediction and check the prediction accuracy  
  
### check the pair plot    
```{r}
pairs(CarbonEmission ~ Body.Type + Sex + Diet + How.Often.Shower + 
    Heating.Energy.Source + Transport + Vehicle.Type + Social.Activity + 
    Monthly.Grocery.Bill + Frequency.of.Traveling.by.Air + Vehicle.Monthly.Distance.Km + 
    Waste.Bag.Size + Waste.Bag.Weekly.Count + How.Long.TV.PC.Daily.Hour + 
    How.Many.New.Clothes.Monthly + How.Long.Internet.Daily.Hour + 
    Energy.efficiency,data = train)
```


```{r}
library(MASS)
stepAIC(waste.fit,scope = list(upper = ~Eth*Sex*Age*Lrn, lower = ~1))
```


```{r}
vars <- step(lm(CarbonEmission ~ 
    Monthly.Grocery.Bill + Frequency.of.Traveling.by.Air + Vehicle.Monthly.Distance.Km + 
    Waste.Bag.Size + Waste.Bag.Weekly.Count + How.Long.TV.PC.Daily.Hour + 
    How.Many.New.Clothes.Monthly + How.Long.Internet.Daily.Hour + 
    Energy.efficiency, data = train), direction = "both")
```

```{r}
s = summary(vars)$sigma
plot(vars$fitted.values, vars$residuals)
abline(h = 0, col = "red")
abline(h = c(-2, 2)*s, col = "red")


```

