---
title: "Dana4820_project"
author: "HassanZhang"
date: "`r Sys.Date()`"
output: html_document
---
```{r}
set.seed(123)
carbon <- read.csv("..//data//Carbon Emission.csv", header = T)

split <- sample(c(TRUE, FALSE), nrow(carbon), replace=TRUE, prob=c(0.7, 0.3))
train <- carbon[split, ]
test <- carbon[!split, ]
```

```{r}
str(train)
```

```{r}
train$CarbonEmission.Level = ifelse(train$CarbonEmission > mean(train$CarbonEmission),'High','Low')
table(train$CarbonEmission.Level)
```

# Q1 . Does diet has a significant effect on carbon emission level ???  with lurking variable 'social activity'...

chi-square test / 2 -way table,

Energy.efficiency
```{r}
table1 = xtabs( ~ CarbonEmission.Level + Diet,data = train)
table1
chisq.test(table1)
```
```{r}
xtabs(~ CarbonEmission.Level + Diet + Social.Activity, data=train)
```
```{r}

boxplot(CarbonEmission ~Diet,data =  train)

library(ggplot2)

# Sample ggplot boxplot command
ggplot(data = train, aes(x = Diet, y = CarbonEmission, fill = Diet)) +
  geom_boxplot() +
  scale_fill_manual(values = c("omnivore" = "red", "pescatarian" = "blue", "vegan" = "green", "vegetarian" = "yellow")) +
  labs(title = "Boxplot of Carbon Emission by Diet",
       x = "Diet",
       y = "Carbon Emission",
       fill = "Diet Category") +
  theme_minimal()

```



```{r}

Social.Activity = c('never',     'often', 'sometimes')
for (Social.Activity_level in Social.Activity){
  print(Social.Activity_level)
  t_i = xtabs(~ CarbonEmission.Level  + Diet, Social.Activity==Social.Activity_level, data=train)
  t_i_oddsratio = epitools::oddsratio(t_i, method = "wald", conf = 0.95, correct = FALSE)
  print(t_i_oddsratio)
  print('-----------------------')
  }
```

###Social.Activity == never: 
Since chi.square p-value > 0.05, we fail to reject H0 𝜃=1, the odd of having a high level carbon emission for each diet category are NOT different for people who never going social , we conclude that there is no association between whether having a high level carbon emission and dietary choices.
From confidence interval results, since 1 is in the CI(0.975007 1.431037) we conclude there is not enough evidence to say there is association between Dietary choices and CarbonEmissionLevel.

###Social.Activity == often: 
Since chi.square p-value <0.05, we reject H0 𝜃=1, the odd of having a high level carbon emission for each diet category are different for people who often going social , we conclude that there is association between whether having a high level carbon emission and dietary choices.
From confidence interval results (1.086801 1.591709), we are 95% confident that the odd of XXX for XXX is between XXX XXX than the odd of XXX for XXX for people who often going social.

###Social.Activity == sometimes: 
Since chi.square p-value <0.05, we reject H0 𝜃=1, the odd of having a high level carbon emission for each diet category are different for people who sometimes going social , we conclude that there is association between whether having a high level carbon emission and dietary choices.

#? how to interpret CI(1.109186 1.644661), which group'odd is higher??

```{r}
library(tidyr)

# Step 1: Create a 3-way contingency table
contingency_table <- xtabs(~ CarbonEmission.Level + Diet + Social.Activity, data=train)

# Step 2: Convert the table to a long-format data frame
long_df <- as.data.frame.table(contingency_table)

# Step 3: Reshape the long-format data frame to a wide format
wide_df <- pivot_wider(long_df, names_from = CarbonEmission.Level, values_from = Freq,
                       id_cols = c(Social.Activity, Diet))
wide_df
# Now `wide_df` is a data frame with `Social.Activity` and `Diet` as hierarchical indices
# and different `CarbonEmission.Level` categories as columns.
```


```{r}
# Sample ggplot boxplot command
ggplot(data = subset(train,Social.Activity == 'never'), aes(x = Diet, y = CarbonEmission, fill = Diet)) +
  geom_boxplot() +
  scale_fill_manual(values = c("omnivore" = "red", "pescatarian" = "blue", "vegan" = "green", "vegetarian" = "yellow")) +
  labs(title = "Boxplot of Carbon Emission by Diet for people never party",
       x = "Diet",
       y = "Carbon Emission",
       fill = "Diet Category") +
  theme_minimal()
```

```{r}
ggplot(data = subset(train,Social.Activity == 'sometimes'), aes(x = Diet, y = CarbonEmission, fill = Diet)) +
  geom_boxplot() +
  scale_fill_manual(values = c("omnivore" = "red", "pescatarian" = "blue", "vegan" = "green", "vegetarian" = "yellow")) +
  labs(title = "Boxplot of Carbon Emission by Diet for people sometimes party",
       x = "Diet",
       y = "Carbon Emission",
       fill = "Diet Category") +
  theme_minimal()
```


```{r}
ggplot(data = subset(train,Social.Activity == 'often'), aes(x = Diet, y = CarbonEmission, fill = Diet)) +
  geom_boxplot() +
  scale_fill_manual(values = c("omnivore" = "red", "pescatarian" = "blue", "vegan" = "green", "vegetarian" = "yellow")) +
  labs(title = "Boxplot of Carbon Emission by Diet for people often party",
       x = "Diet",
       y = "Carbon Emission",
       fill = "Diet Category") +
  theme_minimal()
```


## Q2 GLM  , 1000 dollor increases in bill has 12.6% increase in the number of cloth bought

log(pi) = b0 + b1*bill
bi = 1.255e-04

> exp(1.255e-04)
[1] 1.000126

```{r}

cloth.model = glm(How.Many.New.Clothes.Monthly ~ Monthly.Grocery.Bill ,data = train,family = poisson(link = 'log'))
summary(cloth.model)
```
H0: model is adequate
Ha: model is not adequate
```{r}
D = deviance(cloth.model)
D
df = cloth.model$df.residual
df
pchisq(D,df,lower.tail = FALSE)
```

```{r}
library(mosaic)

mosaic::xpchisq(D,df,lower.tail = FALSE)
critical_value <- qchisq(0.95, df)

#points(critical_value, 0.01, col="red", pch=19)


#text(critical_value, 0.02, labels = paste("Critical value:", round(critical_value, 2)), pos = 4, col="red")

```
### use residucal plot to check the prediction accuract
```{r}
plot(cloth.model)
```



# Q3 ordinal 
Social.Activity 
Energy.efficiency

```{r}
social.energy.table = xtabs(~ Social.Activity + Energy.efficiency, data = train)
social.energy.table
```
```{r}
chisq.test(social.energy.table)
```
```{r}
mosaic::xpchisq(7.8064, 4)
```


```{r}
library(vcdExtra)
A=CMHtest(social.energy.table, rscores = c(0,2,1),cscores=c(0,1,2), types="cor")
A
```


```{r}
#Sample correlation R
n=sum(social.energy.table)
n
M=sqrt(A$table[1])
M
R=M/sqrt(n)
R
#P-value for one-sided standard normal with M statistic

P_value=pnorm(M, lower.tail = FALSE)
P_value
```

```{r}
mosaic::xpnorm(0.633,lower.tail = FALSE)
```




